Finding and controling biases in dark matter halo concentrations estimates

### 2016-07-06 paper/ms.tex
rewriting the conclusions

In this paper we used bootstrapping on dark matter halos resolved with a
large number of particles ($\sim 10^5$) in order to quantify biases on
concentration estimates. 
We found that methods commonly used in the literature present a large
bias in the concentration by overestimating it by factors of
$15\%-20\%$ for halos with $200$ particles. 
The scaling of the bias as a function of particle number was found in
two very different kinds of simulations, one cosmological and the
other of an individual high resolution halo, suggesting its universal
nature.

This provided a motivation to introduce a new method to estimate the
concentration that shows a lower bias.
The method is based on the integrated mass profile. 
We also implemented a fitting procedure based on a Markov Chain Monte Carlo
algorithm.
In the bootstrapping experiment this method showed a bias on the order
of $2\%-3\%$ for halos with $200$ particles.

We provide for the first time in the literature a direct way to
measure the bias in the estimated halo concentration at low particle numbers and
a way to use it to make corrections on the estimates.

We also used the N-body simulation to estimate the impact on the mass
concentration relationship.  
Although the three methods are in broad agreement within their
uncertainties there are some noticeable differences.  
The first difference is that the density method produces
systematically higher concentrations by a factor of $15\%$ compared to
the velocity method.
This systematic offset has been reported in different datasets and
techniques to measure concentrations, without any conclusive
explanation for its origin \citep{Klypin2016}.
The second difference is that the velocity and integrated mass methods
start to differ for masses below $10^{12}\hMsun$.  
We found that a correction of the results by the bias factor found
using the same N-body data can bring these two methods into
agreement.

>>>>

In this paper we quantified biases on concentration numerical estimates by
bootstrapping dark matter halos resolved with a  large number of
particles ($\sim 10^5$).
We found that methods commonly used in the literature present a large
bias in the concentration by overestimating it by factors of
$15\%$-$20\%$ for halos with $200$ particles, or $6\%$-$8\%$ for halos
with $500$ particles. 
This procedure is a new way to quantify the bias in concentration
estimates.

The scaling of the bias as a function of particle number was found using
two very different kinds of simulations, one cosmological and the
other of an individual high resolution halo.

This results motivated the introduction of a new method to estimate the
concentration with a lower bias at low particle numbers.
The method we presented is based on the integrated mass profile. 
To show its performance on simulation data Ww also implemented a
fitting procedure based on a Markov Chain Monte Carlo algorithm.
In the bootstrapping experiment this method showed a bias on the order
of $2\%-3\%$ for halos with $200$ particles and less than $1\%$ for
halos with $500$ particles or more.

We also used the N-body simulation to estimate the impact on the mass
concentration relationship of the new method to estimate
concentration.   
Although the three methods are in broad agreement within their
uncertainties there are some noticeable differences.  
The first difference is that the density method produces
systematically higher concentrations by a factor of $15\%$ compared to
the velocity method.
This systematic offset has been reported in different datasets and
techniques to measure concentrations, without any conclusive
explanation for its origin \citep{Klypin2016}.
The second difference is that the velocity and integrated mass methods
start to differ for masses below $10^{12}\hMsun$.  
We found that a correction of the results by the bias factor found
using the same N-body data can bring these two methods into
agreement.
